---
title: "Exploratory Data Analysis"
format: 
  html:
    code-fold : true
editor: visual
author: Matteo Larrode
execute: 
  cache: true
---

# Salience & Nascence of energy policy areas

## Salience

```{r echo=FALSE, output=FALSE}
#load libraries & packages
library(dplyr)
library(readr)
library(tidyr)
library(ggplot2)
library(ggthemes)
library(lubridate)

salience_df <- read_csv("datasets/salience.csv", show_col_types = FALSE)
salience_df <- salience_df %>% 
  mutate(Subtopic=recode(Subtopic, 
                         'Renewable'='renewable_energy', 
                         'Fossil'='fossil_fuels',
                         'Nuclear'='nuclear_energy',
                         'Solar'= 'solar_power',
                         'Wind'='wind_power',
                         'Hydropower' = 'hydropower',
                         'Biofuels'='biofuels',
                         'Evehicles'='electric_vehicles'))
#turn years into date format (just in case)
salience_df$Year_Date <- lubridate::ymd(salience_df$Year, truncated = 2L)


dataframe_final <- read_csv("datasets/energy_policy_lobbying.csv", show_col_types = FALSE)
```

First, here are representations of the number of times a subtopic is mentioned in a poll question over the years. This is our proxy variable to *salience*.

```{r}
ggplot(salience_df, aes(x=Year))+
  geom_line(aes(y=NbrQuestions, col=Subtopic))+
  labs(title="Salience of energy policing areas (2011-2021)",
       y = "Number of interview questions",
       caption = "Source: Roper iPoll")+
  scale_x_continuous(breaks = seq(from = 2011, to = 2021, by = 2))+
  scale_color_gdocs()+
  theme_solarized()
```

Measuring distinct studies instead of questions allows to avoid inflating salience in the case of a single study asking multiple questions about the same issue.

```{r}

ggplot(salience_df, aes(x=Year))+
  geom_line(aes(y=NbrStudies, col=Subtopic))+
  labs(title="Salience of energy policing areas (2011-2021)",
       y = "Number of Individual Studies",
       caption = "Source: Roper iPoll")+
  scale_x_continuous(breaks = seq(from = 2011, to = 2021, by = 2))+
  scale_color_gdocs()+
  theme_solarized()
```

## Nascence (TO DO)

A first method to measure nascence is to measure the number of years between now and the peak year.

CODE TO DO

A second method is to count the number of years since the subtopic reached a threshold of attention. Here this threshold is set to 5% of the total mentions of energy policy.

CODE TO DO

# Lobbying in energy policy areas

Let's first look at the total number of times each subtopic has been lobbied for or against in the last 10 years, and how much was spent on lobbying

```{r warning=FALSE}
#New dataframe: for each subtopic, global salience & total times lobbied
grouped_areas_df <- dataframe_final %>%
  #get variables of interest
  select(times_lobbied, total_amout, policy_issue) %>%
  #bring to one row per subtopic (summing times lobbied & total amount)
  group_by(policy_issue) %>%
  summarise_all(funs(sum))

ggplot(data = grouped_areas_df, aes(x=policy_issue, y = times_lobbied))+
  geom_bar(stat = "identity", position = "dodge") +
  theme_solarized()+
  ggtitle("Times an energy policy areas was lobbied for or against")+
  theme(axis.text.x = element_text(angle = 65, vjust = 1, hjust=1))

ggplot(data = grouped_areas_df, aes(x=policy_issue, y = total_amout))+
  geom_bar(stat = "identity", position = "dodge") +
  ggtitle("Amount spent by interest groups lobbying in energy policy areas")+
  theme_solarized()+
  theme(axis.text.x = element_text(angle = 65, vjust = 1, hjust=1))
```

We clearly see a perfect match between the amount of money spent on different energy policy areas, and on the amount of times they are lobbied for or against.

# Relationship between salience & lobbying

First, let us look whether or not salience is related to the number of times an issue is lobbied for, and on the money spent.

```{r warning=FALSE}
#collapse salience df to one row per subtopic (summing too)
salience_reduced <- salience_df %>% 
  select(-Year, -Year_Date)%>%
  group_by(Subtopic)%>%
  summarise_all(funs(sum))

#join the two datasets by subtopic
final_df <- left_join(grouped_areas_df, salience_reduced, by = c("policy_issue"="Subtopic"))

```

Looking at the dataframe, and sorting it by order of salience, we notice that it does NOT seem correlated to spending or activity.

Next, it would be interesting to look at whether spending & lobbying activity vary with salience.

TO DO: web scrape the time-series of lobbying activity from 2011-2021 (will be VERY long & heavy)
